---
title: "del_dauer_analysis"
output: html_document
---
```{r echo = FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(igraph)
```


##Load up data and Fucntions for analysis

###Load Variants and Dauer Genes + Filter for Variants in dauer genes 
```{r}
flat_file <- data.table::fread("WI.20210121.strain.annotation-GB-gene-impact-divergent-tempfix.tsv") 
dauer_pathway <- data.table::fread("insulin_genes.tsv")
WB_gene_list <- data.table::fread("wormbase_name_key.txt")
  sample <- sample_n(WB_gene_list, 14)
d_genes <- dauer_pathway$ens_gene
random_genes <- sample$`WormBase Gene ID`

dauer_var <- filter(flat_file, WORMBASE_ID %in% d_genes)
random_var <- filter(flat_file, WORMBASE_ID %in% random_genes)
```

###Load function to do analysis - Outputs Chi Squared Contingency Table, Pairs Table, Pairs Plot

```{r}
var_distribution <- function(df, consequences){ 
#Data Transformations  
  filtered <- df %>% filter(CONSEQUENCE %in% consequences)
  
  strain_split <- filtered %>% separate_rows(Strains, sep = ",") 
  
  strain_group <- strain_split %>% group_by(Strains) %>%  
    summarise(genes_aff = (GENE)) 
  
  contingency <- table(strain_group$Strains,
                       strain_group$genes_aff) %>% as.data.frame.matrix() 
#Chi-Squared Tests 

  chi_table <- cbind(contingency, t(apply(contingency, 1, function(x) {
    ch <- chisq.test(x)
    c(unname(ch$statistic), ch$p.value)})))
  #colnames(df1)[15:16] <- c('x-squared', 'p-value') #The number of columns will change 
  #Add Some formatting?
  
#Chi-Squared Formatting

  

#Looking at Pairings 

logical <- function(x){ ifelse( x > 0, 1 , 0) }
logical_cont <- mutate_all(contingency, logical) #%>%  
logical_cont$genes_affected <- rowSums(logical_cont)
pairs <- filter(logical_cont, genes_affected > 1)

#pairs <-  mutate(sum = rowSums(logical_cont))
 #%>% filter(sum > 1)
  
#Heatmap of pairing by strains
pairs_plot <- pairs %>% as.data.frame() %>% select(-genes_affected) %>% 
  rownames_to_column() %>% 
  gather(Column, Value, -rowname) 
  
plot <- ggplot(pairs_plot, aes(x = rowname, y = Column, fill = Value)) +
  geom_tile() +
  # scale_fill_gradientn(name = "", colors = terrain.colors(3)) +
  scale_x_discrete(name = "") +
  scale_y_discrete(name = "")  

#Co-Occurence Matrix 

melted_data <- melt(strain_group) #what does melt do? 
w <- dcast(melted_data , genes_aff~Strains) # Make data go wider by gene's affected
x <- as.matrix(w[,-1]) #Convert to a matrx 
x[is.na(x)] <- 0
x <- apply(x, 2,  function(x) as.numeric(x > 0)) 
v_f <- x %*% t(x)
diag(v_f) <- 0  #Set Diagonal to Zero
dimnames(v_f) <- list(w[, 1], w[,1])
v_f


longData_f <- melt(v_f)
longData_f<-longData_f[longData_f$value!=0,]

v_plot <- ggplot(longData_f, aes(x = Var2, y = Var1)) + 
  geom_tile(aes(fill=value)) + geom_text(aes(label = value))





# melted_data <- melt(strain_group) #what does melt do? 
# w <- dcast(melted_data, genes_aff~Strains) # Make data go wider by gene's affected
# x <- as.matrix(w[,-1]) #Convert to a matrx 
# x[is.na(x)] <- 0
# x <- apply(x, 2,  function(x) as.numeric(x > 0)) 
# co_table <- x %*% t(x)
# diag(co_table) <- 0  #Set Diagonal to Zero
# dimnames(co_table) <- list(w[, 1], w[,1])
# 
# co_table <- co_table %>% as.data.frame.matrix()


#Plot Co-occurence matrix

# longData <- melt(co_table)
# longData<-longData[longData$value!=0,]
# 
# co_plot <- ggplot(longData, aes(x = Var2, y = Var1)) + 
#   geom_tile(aes(fill=value)) + geom_text(aes(label = value))


#List for outputs 
out <- list()
  out$chi_table <- chi_table 
  out$paired_variants <- pairs
  out$plot <- plot
  out$co_plot <- v_f
  out$co_occurence <- v_plot
  
 
  return(out)

}
```

###Create Consequence Vectors - may need to be updated if there is a compound consequence in the data set
```{r}
high_impact <- c("stop_gain", "start_lost", "splice_donor", "splice_acceptor", "frameshift", "stop_lost")
inframe_altering <- c( "inframe_deletion", "inframe_insertion",  "missense&inframe_altering")
moderate_impact <- c("missense", "moderate_deletion", "moderate_insertion", "missense&inframe_altering", "splice_region")
low_impact <- c("start_retained", "synonymous", "splice_region", "stop_retained")
```


##General Dauer Var Stats


#High Impact Variants

```{r warning = FALSE}
high_analysis <- var_distribution(dauer_var, high_impact) 
high_analysis$chi_table 
high_analysis$paired_variants
high_analysis$plot
high_analysis$co_occurence
high_analysis$co_plot
```


#Inframe Altering Impact Variants

```{r}
inframe_analysis <- var_distribution(dauer_var, inframe_altering) 
inframe_analysis$chi_table 
inframe_analysis$paired_variants
inframe_analysis$plot
inframe_analysis$co_occurence
inframe_analysis$co_plot
```

#High and Inframe Altering

```{r}
high_inframe <- c(high_impact , inframe_altering)
high_inframe_analysis <- var_distribution(dauer_var, high_inframe) 
high_inframe_analysis$chi_table 
high_inframe_analysis$paired_variants
high_inframe_analysis$plot
high_inframe_analysis$co_occurence
high_inframe_analysis$co_plot
```




#Moderate Impact Variants 
```{r}
moderate_analysis <- var_distribution(dauer_var, moderate_impact) 
moderate_analysis$chi_table 
moderate_analysis$paired_variants
moderate_analysis$plot
moderate_analysis$co_occurence
moderate_analysis$co_plot
```



##Missense BSCORE <-2 - There are no -4 Bscore Vars in Dauer Pathway
```{r}
bscore_dauer_var <- dauer_var %>% filter(BLOSUM <= -2)
bscore_analysis <- var_distribution(bscore_dauer_var, moderate_impact)
bscore_analysis$chi_table
bscore_analysis$paired_variants
bscore_analysis$plot
bscore_analysis$co_occurence
bscore_analysis$co_plot
```



##Random Moderate Impact
###There are no High or inframe altering variants in the random data sets
```{r}
random_high_inframe <- var_distribution(random_var, moderate_impact)
random_high_inframe$chi_table
random_high_inframe$paired_variants
random_high_inframe$plot
random_high_inframe$co_occurence
random_high_inframe$co_plot
```



##Classification of Del Variants

```{r}

hi_inframe_b2 <- dauer_var %>% filter(CONSEQUENCE == "missense" & BLOSUM <= -2 | CONSEQUENCE %in% inframe_altering | CONSEQUENCE %in% high_impact ) 
all_con <- c(hi_inframe_b2$CONSEQUENCE) 
hi_inframe_b2_analysis <- var_distribution(hi_inframe_b2, all_con)
hi_inframe_b2_analysis$chi_table
hi_inframe_b2_analysis$paired_variants
hi_inframe_b2_analysis$plot
hi_inframe_b2_analysis$co_occurence
hi_inframe_b2_analysis$co_plot






```




```{r}
hi_inframe_b2_random <- random_var %>% filter(CONSEQUENCE == "missense" & BLOSUM <= -2 | CONSEQUENCE %in% inframe_altering | CONSEQUENCE %in% high_impact ) 
all_con <- c(hi_inframe_b2_random$CONSEQUENCE) 
hi_inframe_b2_random <- var_distribution(hi_inframe_b2_random, all_con)
hi_inframe_b2_random$chi_table
hi_inframe_b2_random$paired_variants
hi_inframe_b2_random$plot
hi_inframe_b2_random$co_occurence
hi_inframe_b2_random$co_plot
```










